### 2016-09-01 main.tex
More minor wording changes.

\begin{abstract}
  The song of the adult male zebra finch is strikingly stereotyped.
  Efforts to understand motor output, pattern generation, and learning
  have taken advantage of this consistency by investigating the bird's
  ability to modify specific parts of song under external cues, and by
  examining timing relationships between neural activity and vocal
  output.  Such experiments require that precise moments during song
  be identified in real time as the bird sings.  Various
  syllable-detection methods exist, but many require special hardware,
  software, and know-how, and details on their implementation and
  performance are scarce.  We present an accurate, versatile, and fast
  syllable detector that can control hardware at precisely timed
  moments during zebra finch song.  Many moments during song can be
  isolated and detected with false negative and false positive rates
  well under 1\% and 0.001\% respectively. The detector can run on a
  stock Mac Mini with triggering delay of less than a millisecond and
  a jitter of $\sigma\approx 2$ milliseconds.
\end{abstract}

@@

This consistency allows a variety of experiments if precise moments in
song can reliably be detected quickly enough to trigger other
apparatus during singing.  A common area of study with this
song-triggering technique is the anterior forebrain pathway (AFP), a
homologue of mammalian basal ganglia consisting of a few distinct
brain areas concerned with the acquisition and learning of song.  For
example, \cite{Kao2005} stimulated the lateral magnocellular nucleus
of the anterior nidopallium (LMAN)---the output nucleus of the
AFP---at precisely timed moments during song and showed that this area
repeatably controls specific variables in song output.
\cite{Wang2008interhemispheric} stimulated LMAN and the high vocal
centre (HVC) in one hemisphere or the other, and found that control of
song rapidly switches between hemispheres.  \cite{Keller2009}
investigated the response of cells in field L and the caudolateral
mesopallium during song with and without feedback perturbations as
well as external sounds, and concluded that these regions may hold a
representation of song against which auditory signals are compared.
\cite{Tumer2007pitchshift} played a disruptive sound during renditions
of a syllable that were slightly above (or below) its average pitch,
and showed that the presence of this apparently random natural
variability in songbird motor output is used to drive change in the
song.  \cite{Andalman2009} extended this experiment and showed that
AFP produces a corrective signal to bias song away from those
disruptions.  \cite{Charlesworth2011learning} showed that the song
change was isolated to within roughly 10 milliseconds (ms) of the
stimulus, and further proposed an elegant model that predicts the
shape of the learned response and suggests a simple mechanism.
\cite{Warren2011} showed that that AFP transfers the error signal to
the robust nucleus of the arcopallium (RA) using
NMDA-receptor--mediated glutamatergic transmission. By looking at song
recovery after applying the same pitch-shift paradigm,
\cite{Canopoli2014} showed that the caudal medial nidopallium is
implicated in memorising or recalling a recent song target, but for
neither auditory processing nor directed motor learning.

@@

Despite the power and versatility of such experiments, there is no standard syllable detector.  Considerations include:
\begin{description}
\item[Accuracy:] How often does the system produce false positives or false negatives?
\item[Latency:] The average delay between the target syllable being sung and the detection.
\item[Jitter:] The amount that latency changes from instance to instance of song.  Our measure of jitter is the standard deviation $\sigma$ of latency.
\item[Versatility:] Is detection possible at ``difficult'' syllables?
\item[Ease of use:] How automated is the process of programming a detector?
\item[Cost:] What are the hardware and software requirements?
\end{description}

A variety of syllable-triggering systems have been used, but few have
been documented or characterised in detail.  In 1999,
\cite{Leonardo1999} used groups of IIR filters with hand-tuned logical
operators.  Their system had a latency of 50 or 100 ms,
and they do not report on jitter or accuracy.  As access to
computational resources has improved, approaches have changed: in
2009, \cite{Andalman2009} still used hand-tuned filters, but ran them
on a Tucker-Davis Technologies digital signal processor.  They report
a latency of around 4 ms, but as with other filter-bank techniques, it
is not strictly a syllable detector but rather a pitch and timbre
detector---it cannot identify a frequency sweep, or distinguish a
short chirp from a long one---and thus requires careful selection of
target syllables.  Furthermore, the method is neither inexpensive nor,
based on our experience with a similar technique, accurate.
\cite{Keller2009} applied a neural network to a spectral image of
song.  They report a jitter of 4.3 ms, but further implementation and
performance details are not available.  In 2011, \cite{Warren2011}
matched spectral templates to stable portions of syllables in 8-ms
segments.  They report a jitter of 4.5 ms, and false-negative and
false-positive rates of up to 2\% and 4\%, respectively.  Hardware
requirements and ease of use were not reported.  In 2013,
\cite{Skocik2013} described in detail a system that matches spectral
images of template syllables using a correlation coefficient.  With a
fast desktop (Intel i7 six-core) running Linux and equipped with a
National Instruments data acquisition card, it boasts a hardware-only
(not accounting for the time taken to compute a match with a syllable)
latency and jitter of just a few microseconds, and the detection
computation they use should not much increase that.  Drawbacks are
that some hand-tuning may still be required, and that they report
false-negative rates around 4\% and 7\% for zebra finches and
Bengalese finches respectively, measured on a small dataset.  In much
other work, an allusion is made to a syllable detector, but no further
information is provided.

@@

\begin{enumerate}
\item Record and align a corpus of training songs.  The technique has been published in \cite{Poole2012}.
\item Choose one or more instants in the song that should create a trigger event, and train a
  neural network to recognise them. This step is carried out offline.  While any neural network software would produce equivalent results, we used MATLAB 2015b's neural network toolbox.
\item Once trained and saved, the neural network is used by a realtime
  detection program that listens to an audio signal and indicates detection of the target
  syllables via a TTL pulse.  We present three implementations, in MATLAB, LabVIEW, and Swift, that trade off hardware requirements, ease of maintenance, and performance.
\end{enumerate}
This method makes the following contributions:
\begin{itemize}
\item Fast: sub-millisecond latencies, with jitter around 2 ms.
\item Accurate: false positive rates under 0.001\% and false negative rates under 1\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
\item State-of-the-art performance with default parameters.
\item Runs on inexpensive hardware.
\item Described in detail here, with reference implementations provided and benchmarked.
\end{itemize}

@@

We begin with a few hundred recordings of a given bird's song, as well as calls, cage noise, and other non-song audio data.  Male
zebra finch song is most highly stereotyped when a female is present
(``directed song'') and slightly more variable otherwise
(``undirected''); we train and test on undirected song since this is
more challenging.  Recordings were made inside a sound-isolating
chamber in which was mounted a recording microphone (Audio-Technica
AT803).  The songs were time-aligned as described in \cite{Poole2012}.

@@

\begin{figure}
  % Width: 8, Height: 3, Fonts scale: 120%
  \includegraphics[width=\textwidth]{Fig1}
  \caption{This image was made by superposing the spectra of our 2818
    aligned songs.  Our example detection points, $t^*_1\ldots t^*_6$,
    are shown as red lines, with the recognition regions (30 ms
    $\times$ 1--8 kHz) marked as rectangles.}
  \label{fig:song}
\end{figure}

@@

With inputs well outside the space on which a neural network has been
trained, its outputs will be essentially random. In order to reduce
the false positive rate it is necessary to provide negative training
examples that include silence, cage noise, wing flapping, non-song
vocalisations, and perhaps songs from other birds.  Although it will
depend on the makeup of the non-song data, we have found that training
with as low as a 1:1 ratio of non-song to song yields excellent
results on most birds.  These data are also included in our accuracy tests.

@@

% technically, it is LabVIEW
\subsubsection{LabVIEW}

@@

The most versatile implementation, MATLAB runs on a variety of
hardware and operating systems, and is perhaps the easiest to modify.
While it did not perform as well on our test system as the other
implementations, the convenience may outweigh the slight timing
performance penalty for some experiments.  Key to minimising jitter is
the size of the audio buffer: on a 2014 Mac Mini running MATLAB 2015b
the smallest buffer size that did not result in read overruns was
about 4 ms.

Similar to the Swift implementation, it is also possible to modify the
MATLAB implementation to generate the TTL pulse from a microcontroller
(Arduino) controlled via a serial over USB interface. This eliminates
the double buffer required when sending triggers via the audio
interface, reducing latency by half.

@@

We have used this detector for song alignment and feedback experiments
in 5 behaving birds, and for offline analysis on a few others.  For
simplicity, and to test robustness to imperfect choice of ``syllable'', we present performance on 6 randomly chosen points in one
bird's song.  However, the results presented here are representative of
real-world performance over many birds in our lab.

@@

The MATLAB neural network toolbox further divides our ``training'' set
into internal training, validation, and test sets.  Because we do our
own testing on a distinct data set, we instructed the
toolbox to use 80\% of its input set for training, and the
remaining 20\% for validation, with no hold-out test data.

@@

We define the accuracy of the network based on its classification
performance per frame. In order to avoid the apparent problem of every
non-detected non-syllable counting as a true negative, we also tried
defining accuracy on a per-song basis, such that a song without the
target syllable counted as a single true negave.  Computing the
optimal output thresholds on a per-frame rather than a per-song basis resulted in higher
thresholds and thus a slight reduction in both the false-positive and true-positive rates (reversible by increasing $C_n$), while also providing a valid
definition of false-positive rate for data streams that had not been
segmented into song-sized chunks and thus allowing easy interpretation
of false-positive rates from calls, cage noise, etc.

@@

Mean latency is then given as the average time between the canonical
signal and the corresponding detection event.  It is a helpful number,
but not a critical one, since in zebra finch a detector with high but
constant latency can often be trained to trigger at a point somewhat
before the true moment of interest (for example, if triggering should
occur at the very beginning of a syllable, the detector's time window
may include previous syllable).  Often more important is the
variability in the latency?  We define jitter as the standard
deviation of the latency.

@@

@@

An example of the detector's output for three of these syllables is
shown in \fig{fig:detection_raster}.  The total audio energy of each
song is shown as a single grayscale row.  By sorting according to time
of each syllable's detection, the rest of the song is shown in the
context of the timing for that moment.  This gives an intuition of the
timing variability within each song and across songs (which may be
responsible for some of our measured jitter).

\begin{figure}
  \includegraphics[width=\textwidth]{Fig2}
  % 9x10 inches
  % Fonts: Helvetica, scaled 130%
  \caption{Each plot shows one network output unit's responses to all
    2818 presentations of the song shown in \fig{fig:song}.  For
    simplicity, we show only the syllables $t^*_1$, $t^*_4$, and
    $t^*_6$, and we do not show the non-response to presentation of non-song. The horizontal axis is time relative to the beginning of
    the aligned song, and the vertical axis is an index for the 2818
    individual song presentations. The grey shading shows the audio
    amplitude of song Y at time T. Detection events on training songs
    are shown in cyan, with detections of unseen test songs in red.  To provide an intuition of intra-song variability, songs have been
    stably sorted by the time of detection events; thus, each of the
    three detection graphs shows the songs in a different order.}
  \label{fig:detection_raster}
\end{figure}

@@

We allowed our training software to allocate our default of 4 hidden
units per syllable, and computed a new FFT every 1.5 ms.  Because our
timing test files are designed for our stereo playback software,
allowing only one channel for the ground-truth pulse, we trained one
detector for each syllable for the following results.  In order to
eliminate the large number of variables involved in microphone, cage,
playback and re-digitising, we evaluated the neural network's accuracy
directly on the digitised recording.  When training and runtime data
are gathered on the same hardware setup, this is the digital signal
that the detector will see---only the non-song data may be different.
%Accuracy \marginpar{Should be! Verify! And/or explain that there is a bug in the LabVIEW implementation, if there in fact is. ALSO need to present a picture of the test syllable or etc\dots} is identical across the three detectors.

\noprint{
  The means of the false positive and

  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_1$
    & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.34\% & 0.0015\%\\ 
    neg & 0.0355\% & 99.99970\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_2$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.92903\% & 0.000596\%\\ 
    neg & 0.0710\% & 99.99940\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_3$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.68062\% & 0.00194\%\\ 
    neg & 0.319\% & 99.99806\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_4$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.82257\% & 0.00239\%\\ 
    neg & 0.177\% & 99.99761\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_5$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.78708\% & 0\\ 
    neg & 0.213\% & 100\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_6$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.71611\% & 0.00552\%\\ 
    neg & 0.284\% & 99.99448\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
}

Accuracies for our six test syllables are shown in
\fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
songs were always 100\%).  For each test syllable we trained
100 different
networks, which differ from each other in the random initialisation of
network weights and on which random subset of our songs was used for
training.  Each network's accuracy is shown as a single point in each
chart in \fig{fig:accuracies}, and the means are shown in Table~\ref{table:accuracies}.  Evidently some syllables are easier to identify
reliably than others; thus, if the experimental design permits, it can be worthwhile to choose an easily identifiable syllable.  Furthermore, as can be seen from \fig{fig:accuracies}, for difficult syllables the training process
occasionally yields a network that performs unusually poorly.  It
is easy to spot these networks by their performance either as measured
as part of the training process or on the test audio file, and
re-train when necessary.

@@

  \caption{Accuracy variability over 100 different training runs for
    each of the six test syllables.  Each dot shows the test-set
    accuracy for an independently trained detector.  Because the
    horizontal positions have been randomised slightly so as not to
    occlude same-valued measurements, test syllable is also indicated
    by colour.  The means are given in Table~\ref{table:accuracies}.}
  \label{fig:accuracies}
\end{figure}

\begin{table}
  \begin{tabular}{c|rr}
    Target time & \% True positives & \% False positives \\ \hline
    $t^*_1$ & 99.34 &  0.0015 \\
    $t^*_2$ & 99.65 &  0.00040 \\
    $t^*_3$ & 98.98 &  0.0015 \\
    $t^*_4$ & 99.30 &  0.00093 \\
    $t^*_5$ & 99.73 &  0.000094 \\
    $t^*_6$ & 98.31 &  0.0026
  \end{tabular}
  \caption{Mean values for the detection accuracies shown in \fig{fig:accuracies}.}
  \label{table:accuracies}
\end{table}

@@

This syllable detector is appropriate for zebra finch song, and
although our tests were carried out on songs from that species, it is
also likely to work well for Bengalese finches due to their similarly stereotyped syllables.  It offers the
following benefits:
\begin{itemize}
\item The detector is accurate. False negative and false positive
  rates can be well under 1\% and 0.001\% respectively, and trading
  these two numbers off against each other is through a single
  relative-cost parameter.
\item Latency is generally under a millisecond, with jitter around 2
  ms.
\item Works on a wide range of target syllables using the default
  values described here, generally eliminating the need for
  hand-tuning.
\item Runs fast enough for syllable-modification experiments on
  inexpensive consumer-grade hardware, although we recommend that the
  training phase be run on a fast desktop system with 32 GB of RAM.
\item A single detector can generate different target pulses for
  multiple syllables at almost no additional computational cost during
  runtime, although training time will increase.
\end{itemize}

>>>>

\begin{abstract}
  The song of the adult male zebra finch is strikingly stereotyped.
  Efforts to understand motor output, pattern generation, and learning
  have taken advantage of this consistency by investigating the bird's
  ability to modify specific parts of song under external cues, and by
  examining timing relationships between neural activity and vocal
  output.  Such experiments require that precise moments during song
  be identified in real time as the bird sings.  Various
  syllable-detection methods exist, but many require special hardware,
  software, and know-how, and details on their implementation and
  performance are scarce.  We present an accurate, versatile, and fast
  syllable detector that can control hardware at precisely timed
  moments during zebra finch song.  Many moments during song can be
  isolated and detected with false negative and false positive rates
  well under 0.5\% and 0.01\% respectively. The detector can run on a
  stock Mac Mini with triggering delay of less than a millisecond and
  a jitter of $\sigma\approx 2$ milliseconds.
\end{abstract}

@@

This consistency allows a variety of experiments if precise moments in
song can reliably be detected quickly enough to trigger other
apparatus during singing.  A common area of study with this
song-triggering technique is the anterior forebrain pathway (AFP), a
homologue of mammalian basal ganglia consisting of a few distinct
brain areas concerned with the production and learning of song.  For
example, \cite{Kao2005} stimulated the lateral magnocellular nucleus
of the anterior nidopallium (LMAN)---the output nucleus of the
AFP---at precisely timed moments during song and showed that this area
repeatably controls specific variables in song output.
\cite{Wang2008interhemispheric} stimulated LMAN and the high vocal
centre (HVC) in one hemisphere or the other, and found that control of
song rapidly switches between hemispheres.  \cite{Keller2009}
investigated the response of cells in field L and the caudolateral
mesopallium during song with and without feedback perturbations as
well as external sounds, and concluded that these regions may hold a
representation of song against which auditory signals are compared.
\cite{Tumer2007pitchshift} played a disruptive sound during renditions
of a syllable that were slightly above (or below) its average pitch,
and showed that the presence of this apparently random natural
variability in songbird motor output is used to drive change in the
song.  \cite{Andalman2009} extended this experiment and showed that
AFP produces a corrective signal to bias song away from those
disruptions.  \cite{Charlesworth2011learning} showed that the song
change was isolated to within roughly 10 milliseconds (ms) of the
stimulus, and further proposed an elegant model that predicts the
shape of the learned response and suggests a simple mechanism.
\cite{Warren2011} showed that that AFP transfers the error signal to
the robust nucleus of the arcopallium (RA) using
NMDA-receptor--mediated glutamatergic transmission. By looking at song
recovery after applying the same pitch-shift paradigm,
\cite{Canopoli2014} showed that the caudal medial nidopallium is
implicated in memorising or recalling a recent song target, but for
neither auditory processing nor directed motor learning.

@@

Despite the power and versatility of vocal feedback experiments, there
is no standard syllable detector.  Desiderata include:
\begin{description}
\item[Accuracy:] How often does the system produce false positives or false negatives?
\item[Latency:] The average delay between the target syllable being sung and the detection.
\item[Jitter:] The amount that latency changes from instance to instance of song.  Our measure of jitter is the standard deviation $\sigma$ of latency.
\item[Versatility:] Is detection possible at ``difficult'' syllables?
\item[Ease of use:] How automated is the process of programming a detector?
\item[Cost:] What are the hardware and software requirements?
\end{description}

A variety of syllable-triggering systems have been used, but few have
been documented or characterised in detail.  In 1999,
\cite{Leonardo1999} used groups of IIR filters with hand-tuned logical
operators.  Their system had a latency of 50 or 100 ms,
and they did not report on jitter or accuracy.  As access to
computational resources has improved, approaches have changed: in
2009, \cite{Andalman2009} still used hand-tuned filters, but ran them
on a Tucker-Davis Technologies digital signal processor.  They reported
a latency of around 4 ms, but as with other filter-bank techniques, it
is not strictly a syllable detector but rather a pitch and timbre
detector---it cannot identify a frequency sweep, or distinguish a
short chirp from a long one---and thus requires careful selection of
target syllables.  Furthermore, the method is neither inexpensive nor,
based on our experience with a similar technique, accurate.
\cite{Keller2009} applied a neural network to a spectral image of
song.  They reported a jitter of 4.3 ms, but further implementation and
performance details are not available.  In 2011, \cite{Warren2011}
matched spectral templates to stable portions of syllables in 8-ms
segments.  They reported a jitter of 4.5 ms, and false-negative and
false-positive rates of up to 2\% and 4\%, respectively.  Hardware
requirements and ease of use were not reported.  In 2013,
\cite{Skocik2013} described in detail a system that matches spectral
images of template syllables using a correlation coefficient.  With a
fast desktop (Intel i7 six-core) running Linux and equipped with a
National Instruments data acquisition card, it boasts a hardware-only
(not accounting for the time taken to compute a match with a syllable)
latency and jitter of just a few microseconds, and the detection
computation they use should not much increase that.  They reported
false-negative rates around 4\% and 7\% for zebra finches and
Bengalese finches respectively, measured on a small dataset.  In much
other work, an allusion is made to a syllable detector, but no further
information is provided.

@@

\begin{enumerate}
\item Record and align a corpus of training songs.  The technique has been published in \cite{Poole2012}.
\item Choose one or more instants in the song that should create a trigger event, and train a
  neural network to recognise them. This step is carried out offline.  While any neural network software would produce equivalent results, we used MATLAB 2015b's neural network toolbox.
\item Once trained and saved, the neural network is used by a realtime
  detection program that listens to an audio signal and indicates detection of the target
  syllables via a TTL pulse.  We present three implementations, in MATLAB, LabVIEW, and Swift, that trade off hardware requirements, ease of maintenance, and performance.
\end{enumerate}
This method makes the following contributions:
\begin{itemize}
\item Fast: sub-millisecond latencies, with jitter around 2 ms.
\item Accurate: false positive rates under 0.02\% and false negative rates under 0.5\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
\item State-of-the-art performance with default parameters.
\item Runs on inexpensive hardware.
\item Described in detail here, with reference implementations provided and benchmarked.
\end{itemize}

@@

We begin with a few hundred recordings of a given bird's song, as well as calls, cage noise, and other non-song audio data.  Male
zebra finch song is most highly stereotyped when a female is present
(``directed song'') and slightly more variable otherwise
(``undirected''); we train and test on undirected song since this is
more commonly studied, as well as more challenging.  Recordings were made inside a sound-isolating
chamber in which was mounted a recording microphone (Audio-Technica
AT803).  The songs were time-aligned as described in \cite{Poole2012}.

@@

\begin{figure}
  \includegraphics[width=\textwidth]{Fig1}
  \caption{This image was made by superposing the spectra of our 2818
    aligned songs.  Our example detection points, $t^*_1\ldots t^*_6$,
    are shown as red lines, with the recognition regions (30 ms
    $\times$ 1--8 kHz) marked as rectangles.}
  \label{fig:song}
\end{figure}

@@

With inputs well outside the space on which a neural network has been
trained, its outputs will be essentially random. In order to reduce
the false positive rate it is necessary to provide negative training
examples that include silence, cage noise, wing flapping, non-song
vocalisations, and perhaps songs from other birds.  Although it will
depend on the makeup of the non-song data, we have found that training
with as low as a 1:1 ratio of non-song to song---or roughly 10-20
minutes of non-song audio---yields excellent results on most birds.
These data are also included in our accuracy tests.

@@

\subsubsection{LabVIEW}

@@

The most versatile implementation, MATLAB runs on a variety of
hardware and operating systems, and is perhaps the easiest to modify.
While it did not perform as well on our test system as the other
implementations, the convenience may outweigh the timing
performance penalty for some experiments.  Key to minimising jitter is
the size of the audio buffer: on a 2014 Mac Mini running MATLAB 2015b
the smallest buffer size that did not result in read overruns was
about 4 ms.

As with the Swift detector, it is also possible to modify the
MATLAB implementation to generate the TTL pulse from a microcontroller
(Arduino) controlled via a serial over USB interface. This eliminates
the double buffer required when sending triggers via the audio
interface, reducing latency by half.

@@

@@

The MATLAB neural network toolbox further divides our ``training'' set
into internal training, validation, and test sets.  Because we do our
own testing on a distinct and much larger data set, we instructed the
toolbox to use 80\% of its training set for training, and the
remaining 20\% for validation, with no hold-out test data.

@@

We define the accuracy of the network based on its classification
performance per frame. In order to avoid the apparent problem of every
non-detected non-syllable counting as a true negative, we also tried
defining accuracy on a per-song basis, such that a song-length sample without the
target syllable counted as a single true negative.  Computing the
optimal output thresholds on a per-frame basis resulted in higher
thresholds and thus a lower false-positive rate, while also providing a valid
definition of false-positive rate for data streams that had not been
segmented into song-sized chunks and thus allowing easy interpretation
of false-positive rates from calls, cage noise, etc.

@@

Mean latency is then given as the average time between the canonical
signal and the corresponding detection event.  It is a helpful number,
but not a critical one, since in zebra finch a detector with high but
constant latency can often be trained to trigger at a point somewhat
before the true moment of interest (for example, if triggering should
occur at the very beginning of a syllable, the detector's time window
may include previous syllable).  Often more important is the
variability in the latency.  We define jitter as the standard
deviation of the latency.

@@

We have used this detector for song alignment and feedback experiments
in 5 behaving birds, and for offline analysis on a few others.  For
simplicity we present performance on 6 randomly chosen points in one
bird's song, but the results presented here are representative of
real-world performance in our lab.

@@

An example of the detector's output for three of these syllables is
shown in \fig{fig:detection_raster}.  The total audio energy of each
song is shown as a single grayscale row.  By sorting according to time
of each syllable's detection, the rest of the song is shown in the
context of the timing for that moment.  This gives an intuition of the
timing variability within each song and across songs (which is
responsible for a small amount of our measured jitter, but note that few of the songs are significantly timeshifted).

\begin{figure}
  \includegraphics[width=\textwidth]{Fig2}
  % 9x10 inches
  % Fonts: Helvetica, scaled 130%
  \caption{Each plot shows one network output unit's responses to all
    2818 presentations of the song shown in \fig{fig:song}.  For
    simplicity, we show only the syllables $t^*_1$, $t^*_4$, and
    $t^*_6$. The horizontal axis is time relative to the beginning of
    the aligned song, and the vertical axis is an index for the 2818
    individual song presentations. The grey shading shows the audio
    amplitude of song Y at time T. Detection events on training songs
    are shown in cyan, with detections of unseen test songs in red.  To provide an intuition of intra-song variability, songs have been
    stably sorted by the time of detection events; thus, each of the
    three detection graphs shows the songs in a different order.}
  \label{fig:detection_raster}
\end{figure}

@@

We allowed our training software to allocate our default of 4 hidden
units per syllable, and computed a new FFT every 1.5 ms.  Because our
timing test files are designed for our stereo playback software,
allowing only one channel for the ground-truth pulse, we trained one
detector for each syllable for the following results.  In order to
eliminate the large number of variables involved in microphone, cage,
playback and re-digitising, we evaluated the neural network's accuracy
directly on the digitised recording.  When training and runtime data
are gathered on the same hardware setup, this is the digital signal
that the detector will see---only the non-song data may be different.

Accuracies for our six test syllables are shown in
\fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
songs were always 100\%).  For each test syllable we trained
100 different
networks, which differ from each other in the random initialisation of
network weights and on which random subset of our songs was used for
training.  Each network's accuracy is shown as a single point in each
chart in \fig{fig:accuracies}.  Some syllables are easier to identify
reliably than others, allowing detection point choice to impact the
detector's accuracy.  For difficult syllables, the training process
occasionally yields a network that performs unusually poorly, but it
is easy to spot these networks by their performance either as measured
as part of the training process or on the test audio file, and
re-train when necessary.

@@

  \caption{Accuracy variability over 100 different training runs for each
    of the six test syllables.  Each dot shows the test-set accuracy
    for an independently trained detector.  Because the horizontal
    positions have been randomised slightly so as not to occlude same-valued
    measurements, test syllable is also indicated by colour.}
  \label{fig:accuracies}
\end{figure}

@@

This syllable detector is appropriate for zebra finch song, and
although our tests were carried out on songs from that species, it is
also likely to work well for Bengalese finches due to their similarly stereotyped syllables.  It offers the
following benefits:
\begin{itemize}
\item The detector is accurate. False negative and false positive
  rates can be well under 0.5\% and 0.01\% respectively, and trading
  these two numbers off against each other is through a single
  relative-cost parameter.
\item Latency is generally under a millisecond, with jitter around 2
  ms.
\item Works on a wide range of target syllables using the default
  values described here, generally eliminating the need for
  hand-tuning.
\item Runs fast enough for syllable-modification experiments on
  inexpensive consumer-grade hardware, although we recommend that the
  training phase be run on a fast desktop system with 32 GB of RAM.
\item A single detector can generate different target pulses for
  multiple syllables at almost no additional computational cost during
  runtime, although training time will increase.
\end{itemize}
